---
title: "linear_models"
author: "Tianyou Wang"
date: "2022-12-01"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Import data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type)
```

```{r}
nyc_airbnb %>% 
  ggplot(aes(x=stars, y=price, color=borough)) +
  geom_point()
```

## Model Fitting

The `lm()` function begins with the formula specification – outcome on the left of the `~` and predictors separated by `+` on the right. As we’ll see shortly, interactions between variables can be specified using `*`. You can also specify an **intercept-only** model `(outcome ~ 1)`, a model with **no intercept** `(outcome ~ 0 + ...)`, and a **model using all available predictors** `(outcome ~ .)`.

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Let's look at the results.

```{r}
fit

summary(fit)
summary(fit)$coef
coef(fit)
```

Using `class()` to check. It is a matrix.
```{r}
summary(fit)$coef %>% class()
```

You can also checked the fitted values and residuals.

```{r eval=FALSE}
fitted.values(fit)
residuals(fit)
```

Let's look at the results better.
Note: they are data frames.

```{r}
broom::glance(fit)

broom::tidy(fit) %>% 
knitr::kable(digits = 3)

broom::tidy(fit) %>% 
  select(-std.error, -statistic) %>% 
  mutate(
    term = str_replace(term, "borough", "Brough: ")
    ) %>% 
  knitr::kable(digits = 3)
```


## Be in controls of factors

R will treat categorical (factor) covariates appropriately and predictably: indicator variables are created for each non-reference category and included in your model, and the factor level is treated as the reference. Usually, alphabetically, the first level is chosen as the reference group.

It’s important to note that changing reference categories won’t change “fit” or statistical sigificance, but can affect ease of interpretation.

Now we want to change the reference `borough` category as the one that have the most rentals. We also want the `room_type` to be treated as a factor variable, not in alphabetical order but treating the reference group as whichever one is most common.

```{r}
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))

nyc_airbnb %>% 
  ggplot(aes(x=stars, y=price, color=borough)) +
  geom_point()

fit = lm(price ~ stars + borough, data = nyc_airbnb)

broom::tidy(fit)
broom::glance(fit)
```

See the graph again, we notice the order of `borough` changed. Manhattan has the most rentals, followed by Brooklyn, Queens, and Bronx. Additionally, we fit the linear model again, Manhattan is the reference group now. The estimators changed. However, we check `broom::glance(fit)`, the model r-sq, AIC, BIC etc. stay the same.





